---
title: Benchmarking Maximum Liklihood functions from R
author: M.Amalan
date: '2018-12-04'
slug: benchmarking-maximum-liklihood-functions-from-r
categories:
  - fitODBOD
tags:
  - fitODBOD
  - mle2
  - optim
  - bbmle
  - nlm
  - ucminf
  - nlminb
  - maxLik
  - mle
image:
  caption: ''
  focal_point: ''
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE,comment = NA,fig.height = 10,fig.width = 10,warning = FALSE,message = FALSE)
library(tidyverse)
library(formattable)
library(kableExtra)
```

# Introduction 

When we need to estimate parameters from a discrete distribution or continuous distribution 
or a function we can use the below mentioned R functions. We will be using the technique of maximizing the 
Log Likelihood function or minimizing the Negative Log Likelihood function.  Based on 
this technique we will compare these R functions because it might benefit people who are struggling
to which one to choose. We have 7 functions in total by my knowledge when I was writing this post. 

Using the [fitODBOD](https://cran.r-project.org/package=fitODBOD) package, I will use the Alcohol Consumption data to try 
and model it for the Beta Binomial Distribution, which has two shape parameters to estimate. The process is that
we find values for shape parameters a and b (or $\alpha$ and $\beta$) which will maximize the Log 
Likelihood function of Beta Binomial Distribution or in our case minimize the Negative Log Likelihood 
function of Beta Binomial Distribution.

Above mentioned Beta-Binomial distributions Probability Mass function is denoted as

$$P_{BetaBin}(x)= {n \choose x} \frac{B(\alpha+x,n+\beta-x)}{B(\alpha,\beta)} $$

where  $n=0,1,2,...$, $x=0,1,2,...,n$ and $\alpha,\beta > 0$. Further $x$ is the Binomial Random variable, 
a,b(or $\alpha$, $\beta$) are shape parameters and $n$ is the binomial trial value. Also B($\alpha$,$\beta$) 
is the beta function. In this distribution we have to estimate the values for a and b. 

Further, using the PMF we can construct the Likelihood function for $\Omega_{BB}=(\alpha,\beta)^T$ as given below: 

$$L(\Omega_{BB}|x)=\prod_{i=1}^{N} \binom{n}{x_i} \frac{B(\alpha+x_i,n+\beta-x_i)}{B(\alpha,\beta)}$$

where N is the Number of observations. Then Negative Log Likelihood function is given as 

$$l(\Omega_{BB}|x)=-\sum_{i=1}^{N} log\binom{n}{x_i} + \sum_{i=1}^{N} log(B(\alpha+x_i,n+\beta-x_i)) - Nlog(B(\alpha,\beta))$$

In the package fitODBOD we have the function EstMLEBetaBin which is constructed based on the above 
Negative Log Likelihood function and we will use it. 

We take Log to transform the Likelihood function values into small values, which will simplify the 
computation process and save time. The estimation functions we need to compare will use specific 
mathematical methods to find the most appropriate shape parameter values.

The functions in question are 

1. [optim](https://www.rdocumentation.org/packages/stats/versions/3.5.1/topics/optim)
2. [nlm](https://www.rdocumentation.org/packages/stats/versions/3.5.1/topics/nlm)
3. [nlminb](https://www.rdocumentation.org/packages/stats/versions/3.5.1/topics/nlminb)
4. [ucminf](https://www.rdocumentation.org/packages/ucminf/versions/1.1-4/topics/ucminf)
5. [maxLik](https://www.rdocumentation.org/packages/maxLik/versions/1.3-4/topics/maxLik)
6. [mle](https://www.rdocumentation.org/packages/stats4/versions/3.5.1/topics/mle)
7. [mle2](https://www.rdocumentation.org/packages/bbmle/versions/1.0.20/topics/mle2)

Further, I will focus on the attributes of the above functions considering from which package, Inputs, Outputs, 
Time to complete estimation and Analytical Methods used with the assistance of tables. Alcohol Consumption data 
has two sets of frequency values but only values from week 1 will be used. Below is the the Alcohol Consumption data, where number of observations
is 399 and the Binomial Random variable is a vector of values from zero to seven. 

```{r Data}
library(fitODBOD)
kable(Alcohol_data,"html",align=c('c','c','c')) %>%
  kable_styling(bootstrap_options = c("striped"),font_size = 14,full_width = F) %>%
  row_spec(0,color = "blue") %>%
  column_spec(1,color = "red")
```

## optim Function

[optim](https://stat.ethz.ch/R-manual/R-devel/library/stats/html/00Index.html#O) is the first function in concern.
[Documentation](https://www.rdocumentation.org/packages/stats/versions/3.5.1/topics/optim) 
of the optim function is useful and it indicates that this function is only used on one input situations only. This means our 
EstMLEBetaBin function has to be modified. Reason for this is that only the parameters that should be estimated need to be input values
but our EstMLEBetaBin function has four parameters which are a,b,x(Binomial Random Variable) and freq(corresponding frequency values). 

While using optim function first index refers to $\alpha$ or a and second index refers to $\beta$ or b. Further, we have 
to input the observations or in our case the Binomial random variable values and their respective frequencies. I think it
is inconvenient to modify the EstMLEBetaBin function, because if we want to estimate parameters for different data-sets it would 
become extra work. After modification we have a new function which is foroptim and I can use it for demonstration and comparison.

Below is the code to estimation and going through the outputs. It should be noted that we have to provide 
initial parameter values as an input to the optim function and it is best to provide values in the domain of  
values a and b. 

Here the shape parameters a and b are in the region of greater than zero but less than positive infinity. 
So for the initial parameters of a=0.1 and b=0.2 we are finding parameters which would minimize the Negative
Log Likelihood function of Beta-Binomial distribution.

```{r optim function from stats package,results='hide'}
# new function to facilitate optim criteria
# only one input but has two elements
foroptim<-function(a)
  {
  EstMLEBetaBin(x=Alcohol_data$Days, freq=Alcohol_data$week1,a=a[1],b=a[2])
  }

# optimizing values for a,b using default inputs
optim_answer<-optim(par=c(0.1,0.2),fn=foroptim)

# obtaining class of output
class(optim_answer)

#length of output
length(optim_answer)

# the outputs
optim_answer$par # estimated values for a, b
optim_answer$value # minimized function value 
optim_answer$counts  # see the documentation to understand
optim_answer$convergence # indicates successful completion
optim_answer$message # additional information

# fitting the Beta-Binomial distribution with estimated shape parameter  values
fitBetaBin(Alcohol_data$Days,Alcohol_data$week1,optim_answer$par[1],optim_answer$par[2])
```

So the foroptim function can be used as above and parameters are estimated for $\alpha$ and $\beta$ (or a, b) for the Alcohol Consumption 
data week 1. Further we have scrutinized the outputs.

* package : stats 
* No of Inputs: 7
* Minimum required Inputs : 2
* Class of output : list
* No of outputs: 5
* No of Analytical Methods : 6
* Default Method : Nelder-Mead

## nlm Function

[nlm](https://stat.ethz.ch/R-manual/R-devel/library/stats/html/00Index.html#N) function is also similar to optim 
but only one analytical method will be used, which is a Newton-type Algorithm. 
Here also there need to be changes made to our EstMLEBetaBin function as previously. After making those changes 
we have called the new Negative Log-likelihood R function as fornlm. Then we can use the nlm function and estimate 
a and b which are the initial shape parameter values of 0.1 and 0.2 respectively. 
[Documentation](https://www.rdocumentation.org/packages/stats/versions/3.5.1/topics/nlm) 
of nlm function is very useful to understand how it works.

Below is the code for using nlm function appropriately and fiddling with the results.

```{r nlm function from stats package,results='hide'}
#new function to facilitate nlm criteria
fornlm<-function(a)
  {
  EstMLEBetaBin(x=Alcohol_data$Days, freq=Alcohol_data$week1,a=a[1],a[2])
  }

#optimizing values for a,b using default input
nlm_answer<-nlm(f=fornlm,p=c(0.1,0.2))

#obtaining class of output
class(nlm_answer)

#length of output
length(nlm_answer)

# the outputs
nlm_answer$estimate # estimated values for a, b
nlm_answer$minimum # minimized function value 
nlm_answer$gradient  # gradient at the estimated minimum of given funciton
nlm_answer$code # indicates successful completion
nlm_answer$iterations # number of iterations performed

# fitting the Beta-Binomial distribution with estimated shape parameter  values
fitBetaBin(Alcohol_data$Days,Alcohol_data$week1,nlm_answer$estimate[1],nlm_answer$estimate[2])
```

Similarly for the fornlm function we have estimated values for a and b which would fit the Alcohol consumption data.
of week 1.

* package : stats 
* No of Inputs: 12
* Minimum required Inputs : 2
* Class of output : list
* No of outputs: 5
* No of Analytical Methods : 1
* Default Method : Newton-type Algorithm

## nlminb Function

[nlminb](https://stat.ethz.ch/R-manual/R-devel/library/stats/html/00Index.html#N) is also similar and 
requires that EstMLEBetaBin function to be restructured as previous two situations. 
After this task we now have a function called fornlminb. nlminb function is based on analytical method of 
unconstrained and box-constrained optimization using PORT routines.

After choosing initial parameter values for a and b, which are respectively 0.1 and 0.2 the estimation was done using
nlminb function. 

```{r nlminb function from stats package, results='hide'}
# new function to facilitate nlminb criteria
fornlminb<-function(a)
  {
  EstMLEBetaBin(x=Alcohol_data$Days, freq=Alcohol_data$week1,a=a[1],a[2])
}

# optimizing values for a,b using default inputs
nlminb_answer<-nlminb(start=c(0.1,0.2), objective=fornlminb)

# obtaining class of output
class(nlminb_answer)

# length of output
length(nlminb_answer)

# the outputs
nlminb_answer$par # estimated values for a, b
nlminb_answer$objective # minimized function value 
nlminb_answer$evaluations  # see the documentation to understand
nlminb_answer$convergence # indicates successful completion
nlminb_answer$message # additional information
nlminb_answer$iterations # number of iterations performed

# fitting the Beta-Binomial distribution with estimated shape parameter values
fitBetaBin(Alcohol_data$Days,Alcohol_data$week1,nlminb_answer$par[1],nlminb_answer$par[2])
```

[Documentation](https://www.rdocumentation.org/packages/stats/versions/3.5.1/topics/nlminb) 
includes the information related to the function therefore referring it will 
be useful. After estimating values for a and b using nlminb function these were noticed as before in 
input and output criteria

* package : stats 
* No of Inputs: 8
* Minimum required Inputs : 2
* Class of output : list
* No of outputs: 6
* No of Analytical Methods : 1
* Default Method : Unconstrained and box-constrained optimization using PORT routines

## ucminf Function

Package [uncminf](https://cran.r-project.org/package=ucminf) produces the ucminf function, as previously mentioned functions here also
we have to change the EstMLEBetaBin function. After making the changes we will be using the
forucminf function to the estimation process of shape parameters a and b. 

When initial parameter values are set to a=0.1 and b=0.2 we will obtain results from 
ucminf function, which will minimize the Negative Log-likelihood function. Below is the 
code for estimation and using the results to understand the function ucminf.

```{r ucminf function from stats package, results='hide'}
library(ucminf)

# new function to facilitate ucminf criteria
forucminf<-function(a)
  {
  EstMLEBetaBin(x=Alcohol_data$Days,freq = Alcohol_data$week1,a=a[1],a[2])
}

# optimizing values for a,b using default methods
ucminf_answer<-ucminf(par=c(0.1,0.2),fn=forucminf)

#obtaining class
class(ucminf_answer)

# length of output
length(ucminf_answer)

# the outputs
ucminf_answer$par # estimated values for a, b
ucminf_answer$value # minimized function value 
ucminf_answer$invhessian.lt  # see the documentation understand
ucminf_answer$convergence # indicates successful completion
ucminf_answer$message # additional information
ucminf_answer$info

# fitting the Beta-Binomial distribution with estimated shape parameter values
fitBetaBin(Alcohol_data$Days,Alcohol_data$week1,ucminf_answer$par[1],ucminf_answer$par[2])
```

With the help of R [Documentation](https://www.rdocumentation.org/packages/ucminf/versions/1.1-4/topics/ucminf) the 
estimation was done for shape parameter values a and b using the ucminf function. Below are the initial understandings of
the inputs and outputs of ucminf function 

* package : ucminf
* No of Inputs: 5
* Minimum required Inputs : 2
* Class of output : list
* No of outputs: 6
* No of Analytical Methods : 1
* Default Method : Quasi-Newton Algorithm type with BFGS updating

## maxLik Function

maxLik function is the cause of [maxLik](https://cran.r-project.org/package=maxLik) package, which only 
maximizes the Log Likelihood function. Therefore we have to restructure EstMLEBetaBin as previously mentioned but as 
an addition a negative sign is added for the output. This new function will be called as formaxLik. 

For the initial parameter values where a=0.1 and b=0.2 the maxLik function will be used and 
results will be evaluated as below. 

```{r maxLik function from maxLik package, results='hide'}
library(maxLik)

# new function to facilitate maxLik criteria
formaxLik<-function(a)
  {
  -EstMLEBetaBin(x=Alcohol_data$Days,freq = Alcohol_data$week1,a=a[1],a[2])
  }

# optimizing values for a,b using default methods
maxLik_answer<-maxLik(logLik =formaxLik, start = c(0.1,0.2))

# obtaining class of output
class(maxLik_answer)

# length of output
length(maxLik_answer)

# the outputs
maxLik_answer$estimate # estimated values for a, b
maxLik_answer$maximum # minimized function value 
maxLik_answer$iterations  # no of iterations to succeed
maxLik_answer$gradient # last gradient value which was calculated
maxLik_answer$message # additional information
maxLik_answer$hessian # hessian matrix
maxLik_answer$code # indicates successful completion
maxLik_answer$fixed # logical vector indicating which parameters are constants
maxLik_answer$type # type of maximization
maxLik_answer$last.step # list describing the last unsuccessful step
maxLik_answer$control # see the documentation understand

# fitting the Beta-Binomial distribution with estimated shape parameter values
fitBetaBin(Alcohol_data$Days,Alcohol_data$week1,maxLik_answer$estimate[1],maxLik_answer$estimate[2])
```

Using the [Documentation](https://www.rdocumentation.org/packages/maxLik/versions/1.3-4/topics/maxLik) of maxLik
and further using the [Documentation](https://www.rdocumentation.org/packages/maxLik/versions/1.3-4/topics/maxNR) of 
maxNR function. According to the code above below are the findings in for maxLik function in point form 

* package : maxLik
* No of Inputs: 6
* Minimum required Inputs : 2
* Class of output : list or class of maxim or class of maxLik
* No of outputs: 11
* No of Analytical Methods : 7
* Default Method : Automatically chosen

## mle Function

[stats4](https://stat.ethz.ch/R-manual/R-devel/library/stats4/html/00Index.html) package is installed so that mle function can be operated for the purpose of 
estimating a and b parameters. Here also as previously we need to make some changes 
as below and create a new estimating function called formle. 

For the initial parameter values of a=0.1 and b=0.2 the Negative Log Likelihood value 
has been minimized using mle function. Below is the code for estimation and investigation
from the outputs of mle after estimation. 

```{r mle function from stats4 package, results='hide'}
library(stats4)

# new function to facilitate mle criteria
formle<-function(a,b)
  {
  EstMLEBetaBin(x=Alcohol_data$Days,freq = Alcohol_data$week1,a,b)
  }

# optimizing values for a,b using default methods
mle_answer<-mle(minuslogl=formle,start = list(a=0.1,b=0.2))

# obtaining class
class(mle_answer)

# length of output
length(mle_answer)

# the outputs
mle_answer@call # inputs i have used 
mle_answer@coef # estimated values for a,b
mle_answer@fullcoef # all values , even the fixed values we did not want to estimate
mle_answer@vcov # variance covariance matrix for a,b
mle_answer@min # minimized function value
mle_answer@details # details after estimation process
mle_answer@nobs # number of observations to be used for computing only if given by us
mle_answer@method # optimization methods used

# Methods used
confint(mle_answer) # confidence intervals for estimated values
logLik(mle_answer) # Negative loglikelihood value for estimated values 
profile(mle_answer) # Likelihood profile generation.
nobs(mle_answer)  # number of observations to be used for computing only if given by us
show(mle_answer) # display object briefly
summary(mle_answer) # generate a summary
#update()   # updating if we have new data and need to estimate new values
vcov(mle_answer) # variance covariance matrix

# fitting the Beta-Binomial distribution with estimated shape parameter values
fitBetaBin(Alcohol_data$Days,Alcohol_data$week1,mle_answer@coef[1],mle_answer@coef[2])
```

It should be noted that in all 5 previous function we only saw lists of output, but
through mle we are seeing a class of mle output.The [Documention](https://www.rdocumentation.org/packages/stats4/versions/3.5.1/topics/mle) 
explains the inputs and outputs of mle. [Documentation](https://www.rdocumentation.org/packages/stats4/versions/3.5.1/topics/mle-class) 
for mle-class explains further about the methods that can be used. Further, below is a list of findings 
based on the outputs from mle function.

* package : stats4
* No of Inputs: 5
* Minimum required Inputs : 2
* Class of output : class of mle
* No of outputs: 9
* Methods for output: 8 
* No of Analytical Methods :6
* Default Method : Nelder and Mead

## mle2 Function

mle2 function is advanced than mle function and it is from the package [bbmle](https://cran.r-project.org/package=bbmle). 
Here, there is no need to modify the EstMLEBetabin function from the fitODBOD package 
to estimation as all previous situations . 

For the initial parameter values of a=0.1 and b=0.2 Negative Log Likelihood function will be
minimized where outputs will be investigated and methods related to output in class of mle2 will 
be used.

Below is the code for using mle2 function and scrutinizing the output and methods 
relate to it. 

```{r mle2 function from bbmle package, results='hide'}
library(bbmle)

# optimizing values for a,b using default methods
mle2_answer<-mle2(minuslogl= EstMLEBetaBin,start = list(a=0.1,b=0.2),
                  data = list(x=Alcohol_data$Days,freq=Alcohol_data$week1))

# obtaining class
class(mle2_answer)

# length of output
length(mle2_answer)

# the outputs
mle2_answer@call # inputs generally considered 
mle2_answer@call.orig # inputs i have given
mle2_answer@coef # estimated values for a,b
mle2_answer@fullcoef # all values , even the fixed values we did not want to estimate
mle2_answer@vcov # variance covariance matrix for a,b
mle2_answer@min # minimized function value
mle2_answer@details # details after estimation process
mle2_answer@method # optimization methods used
mle2_answer@data # data used for estimation 
mle2_answer@formula # if a formula was specified in the input 
mle2_answer@optimizer # function used for optimizing

# Methods used
coef(mle2_answer) # extrat the estimated values
confint(mle2_answer) # confidence intervals for estimated values
show(mle2_answer) # display object briefly
summary(mle2_answer) # generate a summary
#update()   #updating if we have new data and need to estimate new values
vcov(mle2_answer) # variance covariance matrix
#formula(mle2_answer) # if a formula was specified in the input 
#plot(mle2_answer) # plot the profile
logLik(mle2_answer) # Negative loglikelihood value for estimated values 
profile(mle2_answer) # profile of estimated values

# fitting the Beta-Binomial distribution with estimated shape parameter values
fitBetaBin(Alcohol_data$Days,Alcohol_data$week1,mle2_answer@coef[1],mle2_answer@coef[2])
```

After estimation, fiddling with the output, with the [Documention](https://www.rdocumentation.org/packages/bbmle/versions/1.0.20/topics/mle2) of
mle2 function and mle2-class [Documentation](https://www.rdocumentation.org/packages/bbmle/versions/1.0.20/topics/mle2-class) few findings are mentioned below
in point form.

* package : bbmle
* No of Inputs: 22
* Minimum required Inputs : 3
* Class of output : class of mle2
* No of outputs: 12
* Methods for output: 9
* No of Analytical Methods : 6
* Default Method : Nelder and Mead

# Summary of package Information of the R functions 

After completing a brief fact finding of the R functions optim, nlm, nlminb, ucminf
maxLik, mle and mle2 we can record the facts and results in a table. Using this
table we can choose the best suitable function for our needs of estimating parameters.

I would prefer mle2 function than others because it provides special methods to handle
the mle2 outputs and the output themselves are very thorough than other R functions. It 
can be seen that even though estimated a and b shape parameter values seem different only 
from the fourth decimal point in between all outputs. But this does not effect the 
minimized Negative Log Likelihood function which is same for all seven function outputs.

```{r Summary of package info,input and output,echo=FALSE}
Results1<-data.frame(
                Function=c("Package","No of Inputs","Minimum No of Inputs","Analytical Methods","Output Type", "No of Outputs","No of Methods",
                "Initial value(a,b)","Estimated value(a,b)","Negative Log Likelihood value"),
                optim=c("stats","7","2","6","List","5","None","a=0.1\nb=0.2","a=0.7230707 \nb=0.5809894","813.4571"),
                nlm=c("stats","12","2","1","List","5","None","a=0.1\nb=0.2","a=0.7229384 \nb=0.5808448","813.4571"),      
                nlminb=c("stats","8","2","1","List","6","None","a=0.1\nb=0.2","a=0.7229404 \nb=0.5808469","813.4571"),  
                ucminf=c("ucminf","5","2","1","List","6","None","a=0.1\nb=0.2","a=0.7229390 \nb=0.5808458","813.4571"),
                maxLik=c("maxLik","6","2","7","List \nclass maxLik \nclass maxim","11","None","a=0.1\nb=0.2","a=0.7229428 \nb=0.5808488","813.4571"),
                mle=c("stats4","5","2","6","class mle","9","8","a=0.1\nb=0.2","a=0.7228930 \nb=0.5807279","813.4571"),
                mle2=c("bbmle","22","3","6","class mle2","12","9","a=0.1\nb=0.2","a=0.7228930 \nb=0.5807279","813.4571")
                    )
kable(Results1,"html",align = c('c','c','c','c','c','c','c','c')) %>%
  kable_styling(full_width = T,bootstrap_options = c("striped"),font_size = 14) %>%
   row_spec(0,color = "blue") %>%
   column_spec(1,color = "red")
```

# Summary of Time evaluation for the R functions

Further at the beginning I have considered to evaluate the time of system process 
for the functions to produce results and compare them. In order to do this time comparison
it is possible to use the [benchmark](https://www.rdocumentation.org/packages/rbenchmark/versions/1.0.0/topics/benchmark) 
function of [rbenchmark](https://cran.r-project.org/package=rbenchmark) 
package and below mentioned code chunk provides the output 
in a table form which includes the functions and their respective time values. The estimation 
process of the parameters where each function has been replicated 1000 times to receive a
more accurate and strong table of time values. 


The table is order according to the elapsed time value column in ascending order. According to this we 
can see that least time takes to the nlminb function and most time is taken to the mle2 function. These 
times complete depends on the Negative Log Likelihood function you need to minimize, the data you provided,
the number of estimators that needs to be estimated and finally the complexity of the function. 

Therefore based on the needs of your output and the function which needs estimation choose the best function.
Because as we say in the earlier table the estimated values are slightly different but does not make any strong influence 
on the results.

```{r time benchmark}
library(rbenchmark)

Results2<-benchmark(
          "optim"={optim(par = c(0.1,0.2), fn = foroptim)},
          "nlm"={nlm(f = fornlm, p = c(0.1,0.2))},
          "nlminb"={nlminb(start = c(0.1,0.2), objective = fornlminb)},
          "ucminf"={ucminf(par = c(0.1,0.2), fn = forucminf)},
          "maxLik"={maxLik(logLik = formaxLik, start = c(0.1,0.2))},
          "mle"={mle(minuslogl = formle, start = list(a=0.1,b=0.2))},
          "mle2"={mle2(minuslogl = EstMLEBetaBin, start = list(a=0.1, b=0.2),
                       data = list(x=Alcohol_data$Days, freq=Alcohol_data$week1))},
          replications = 1000,
          columns = c("test","replications","elapsed",
                      "relative","user.self","sys.self"),
          order = 'elapsed'
          )

kable(Results2,"html",align = c('c','c','c','c','c','c')) %>%
  kable_styling(full_width = T,bootstrap_options = c("striped"),font_size = 14) %>%
  row_spec(0,color = "blue") %>%
  column_spec(1,color = "red")
```

# Summary of Results after estimating parameters using the R functions

After using the functions optim, nlm, nlminb, ucminf, maxLik, mle and mle2 to estimate
the shape parameters a, b we can use the estimated parameters in the function fitBetaBin.
Using this function we can find expected frequencies for each of these functions and 
compare p-values and over-dispersion and understand if using different estimation functions had
any effect on them.

According to the below table there is no significant changes between the expected frequencies,
p-values or the over dispersion values. This is a clear indication of it does not matter what function we
use the estimation will occur effectively but only efficiency will be affected.

```{r Summary of Results,echo=FALSE}
Results3<-data.frame(
BinomialRandomVariable=c("0","1","2","3","4","5","6","7","Total No of Observations","p-value","Over Dispersion"),
               Frequency=c("47","54","43","40","40","41","39","95","399","",""),
                   optim=c("54.61","42","38.91","38.54","40.07","44","53.09","87.77","398.99","0.0902","0.4340165"),
                     nlm=c("54.62","42","38.9","38.54","40.07","44","53.09","87.78","399","0.0901","0.4340686"),
                  nlminb=c("54.62","42","38.9","38.54","40.07","44","53.09","87.78","399","0.0901","0.4340679"),
                  ucminf=c("54.62","42","38.9","38.54","40.07","44","53.09","87.78","399.01","0.0901","0.4340683"),
                  maxLik=c("54.62","42","38.9","38.54","40.07","44","53.09","87.78","399","0.0901","0.434067"),
                     mle=c("54.62","42","38.9","38.54","40.07","43.99","53.09","87.8","399.01","0.0903","0.4340992"),      
                     mle2=c("54.62","42","38.9","38.54","40.07","43.99","53.09","87.8","399.01","0.0903","0.4340992")  
                  )

kable(Results3,"html",align = c('c','c','c','c','c','c','c','c')) %>%
  kable_styling(full_width = T,bootstrap_options = c("striped"),font_size = 14) %>%
  row_spec(0,color = "blue") %>%
  column_spec(1,color = "red")

```

# Final conclusion

We had 7 functions to compare but choosing one over the other is completely harmless
to the final result of estimation as seen by our tables. The only issue is time, therefore
I would recommend chose the best function based on your needs of output and research objective.

*Thank You*
