---
title: 'Week 23 : Ramen Ratings'
author: M.Amalan
date: '2019-06-05'
slug: Week23
output:
    blogdown::html_document:
      toc: true
categories:
  - TidyTuesday
tags:
  - TidyTuesday
  - tidyverse
  - gganimate
image:
  caption: ''
  focal_point: ''
summary: "2019 Week 23 TidyTuesday: Ramen Ratings"  
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE,warning = FALSE,message = FALSE,fig.width = 12,fig.height = 12)
```

```{r load th data and packages}
ramen_ratings <- readr::read_csv("https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2019/2019-06-04/ramen_ratings.csv")

library(tidyverse)
library(gganimate)
library(ggthemes)
```

```{r Stars all}
ramen_ratings %>%
  count(stars,sort = TRUE) %>%
ggplot(.,aes(fct_inorder(as_factor(stars)),n,label=n))+
      geom_col(fill=blues9[5])+
      geom_text(vjust=-0.5,color=blues9[9])+ylab("Count / Frequency")+
      xlab("Stars")+ggtitle("Stars Distribution")+
      theme_economist()+
      theme(axis.text.x = element_text(angle = 45))
```

## Brand versus Stars


### Rating where stars is 4 or above 4

```{r Rating stars of 4 or above 4 for Brands}
ramen_ratings %>%
  subset(stars >=4)%>%
  count(brand,sort = TRUE) %>%
  top_n(25) %>%
ggplot(.,aes(x=fct_inorder(brand),y=n,label=n))+geom_col(fill=blues9[5])+
      ylab("Count / Frequency")+xlab("Brand")+
      geom_text(hjust=-0.5,color=blues9[9])+
      ggtitle("Brands Distribution by Stars",
              subtitle = "Top 25")+
      theme_economist()+ coord_flip()
```

### Rating where stars is below 2

```{r Rating stars of 2 or below 2 for Brands}
ramen_ratings %>%
  subset(stars < 2)%>%
  count(brand,sort = TRUE) %>%
  top_n(25) %>%
ggplot(.,aes(x=fct_inorder(brand),y=n,label=n))+geom_col(fill=blues9[5])+
      ylab("Count / Frequency")+xlab("Brand")+
      geom_text(hjust=-0.5,color=blues9[9])+
      ggtitle("Brands Distribution by Stars",
              subtitle = "Top 25")+
      theme_economist()+ coord_flip()
```

## Style versus Stars

### Rating where stars is 4 or above 4

```{r Rating stars of 4 or above 4 for Style}
ramen_ratings %>%
  subset(stars >=4)%>%
  count(style,sort = TRUE) %>%
  top_n(10) %>%
ggplot(.,aes(x=fct_inorder(style),y=n,label=n))+geom_col(fill=blues9[5])+
      ylab("Count / Frequency")+xlab("Style")+
      geom_text(hjust=-0.5,color=blues9[9])+
      ggtitle("Styles Distribution by Stars",
              subtitle = "Top 10")+
      theme_economist()+ coord_flip()
```


### Rating below 2

```{r Rating stars of 2 or below 2 for Style}
ramen_ratings %>%
  subset(stars < 2)%>%
  count(style,sort = TRUE) %>%
  top_n(5) %>%
ggplot(.,aes(x=fct_inorder(style),y=n,label=n))+geom_col(fill=blues9[5])+
      ylab("Count / Frequency")+xlab("Style")+
      geom_text(hjust=-0.5,color=blues9[9])+
      ggtitle("Styles Distribution by Stars",
              subtitle = "Top 5")+
      theme_economist()+ coord_flip()
```

## Country versus Stars

### Rating where stars is 5

### Rating below 2

## Brand and Style versus Stars

### Rating where stars is 5

### Rating below 2

## Brand and Country versus Stars

### Rating where stars is 5

### Rating below 2

## Style and Country versus Stars

### Rating where stars is 5

### Rating below 2

## Brand, Style and Country versus Stars

### Rating where stars is 5

### Rating below 2