---
title: 'Benchmarking the mle and mle2 function '
author: M.Amalan
date: '2018-12-29'
output: 
  blogdown::html_page:
    toc: true
slug: benchmarking-the-mle-and-mle2-function
categories:
  - R
tags:
  - bbmle
  - fitODBOD
  - mle
  - mle2
  - R
  - R package
image:
  caption: ''
  focal_point: ''
---

# Introduction

[mle](https://www.rdocumentation.org/packages/stats4/versions/3.5.1/topics/mle) and [mle2](https://www.rdocumentation.org/packages/bbmle/versions/1.0.20/topics/mle2) 
are my favorite functions, because they provide extensive amount of outputs
for the optimization process. Even though there is no difference in analytial methods 
used in both of these functions. Further, these analytical methods are the same ones used
by optim function. To be honest mle and mle2 functions are wrapper functions of optim. It means
both mle and mle2 are using the optim function inside but with some additional inputs, which
would generate extended outputs.

Even if I do Benchmark the analytical methods for the mle function it would be very similar to
optim function tables but with additional time taken, because of the extra outputs. This would 
similarly occur when we benchmark analytical methods from the mle2 function as well. 

Therefore, I figure why do we need to benchmark them at all. So this blog post is to 
simply reiterate the initial things which I said in my earlier post on the blog post
[Benchmarking optimization functions in R](https://amalan-con-stat.netlify.com/post/fitodbod_1/benchmarking-maximum-liklihood-functions-from-r/).

## mle

mle function is from the [stats4](https://stat.ethz.ch/R-manual/R-devel/library/stats4/html/00Index.html) 
package. If we intend to use this function for
the estimation of shape parameters a and b of the Beta-Binomial distribution when
Binomial Outcome Data, then we need to use the EstMLEBetaBin function from the fitODBOD
package. This is not enough because for limitations in the mle we need to make changes 
in our EstMLEBetaBin function as mentioned below. 

```{r using mle,warning=FALSE,message=FALSE,results='hide'}
library(stats4)
library(fitODBOD)

#new function to facilitate mle criteria 
formle<-function(a,b)
{
  EstMLEBetaBin(x=Alcohol_data$Days,freq = Alcohol_data$week1,a,b)
}

# optimizing values for a,b using default analytial method
mle_answer<-mle(minuslogl = formle,start = list(a=0.1,b=0.2))
```

We are going to use the Alcohol Consumption data of week 1. In the above code
chunk we are using the mle function for our task of finding the optimum shape parameter
values for a and b while using the give Binomial Outcome data. Also If you wish you
study about the mle function refer [this link](https://amalan-con-stat.netlify.com/post/fitodbod_1/benchmarking-maximum-liklihood-functions-from-r/#mle-function) 
from my previous post.

## mle2

[bbmle](https://cran.r-project.org/web/packages/bbmle/index.html) 
package holds the mle2 function. It is simply an updated version for the mle function. 
Although there need to be no changes in the EstMLEBetaBin function to satisfy the mle2 
function's criteria. Now it will be possible to use it. 

```{r using mle2 function}
library(bbmle)

# optimizing values for a,b using default analytical method
mle2_answer<-mle2(minuslogl= EstMLEBetaBin,start = list(a=0.1,b=0.2),
                  data = list(x=Alcohol_data$Days,freq=Alcohol_data$week1))
```


Still if someone needs a brief introduction to mle2 function they can refer my
previous brief through [this link](https://amalan-con-stat.netlify.com/post/fitodbod_1/benchmarking-maximum-liklihood-functions-from-r/#mle2-function).

# Conclusion

My personal opinion is to use the mle2 function, but moving towards what should be the analytical
method. It would be wise to choose it based on your needs as these methods completely depend on 
the data, function that needs to be estimated, complexity of the function and finally the number
estimators that needs to be estimated. 

This is the link to the article which is for 
[Benchmarking optim function analytical methods](https://amalan-con-stat.netlify.com/post/optim/optim-estimating-the-shape-parameters-of-beta-binomial-distribution/).
It might be useful while understanding the analytical methods.
