---
title: Tree of Binomial Distribution
author: M.Amalan
date: '2019-02-14'
slug: BinomialDistribution
output:
  blogdown::html_page:
    toc: true
categories:
  - fitODBOD
  - Blog
  - Binomi
tags:
  - Beta-Binomial
  - fitODBOD
  - Binomial
image:
  caption: ''
  focal_point: ''
summary: 'Brief introduction on Binomial Distribution and its expansion based on data.'  
---

# Binomial Distribution

The binomial distribution can be defined, using the binomial expansion 

$$ (q+p)^n = \sum_{x=0}^{n} {n \choose k} p^k q^{(n-k)} = \sum_{x=0}^{n} \frac{n!} {k! (n-k)!} p^k q^{(n-k)}$$

as the distribution of a random variable X for which 

$$Pr[X=x] = {n \choose k} p^k q^{(n-k)}$$ 

$x=0,1,2,...,n.$ where $q+p=1$, $p,q>0$ and $n$ is a positive integer. When $n=1$, the distribution is known as the 
Bernoulli distribution. The mean and variance are $\mu=np$ and  $\mu_2 = npq$.

If $n$ independent trials are made and in each there is probability $p$ that the outcome $E$ will occur, 
then the number of trials in which $E$ occurs can be represented by a random variable $X$ having the binomial distribution 
with parameters $n$, $p$. 

This situation occurs when a sample of fixed size $n$ is taken from an infinite population 
where each element in the population has an **"equal"** and **"independent"** probability $p$ of possession of a specified 
attribute. The situation also arises when a sample of fixed size $n$ is taken from a infinite population where each 
element in the population has an **"equal"** and **"independent"** probability $p$ of having a specified attribute 
and elements are sampled independently and sequentially with replacement. 

The distribution was derived by James Bernoulli (in his treatise Ars Conjectandi, published in 1713), 
for the case $p = r/(r+s)$, where $r$ and $s$ are positive integers.

The binomial distribution is of such importance in applied probability and statistics that it is frequently 
necessary to calculate probabilities based on this distribution. Although the calculation of sums of the form

$$ \sum_{x} {n \choose x} p^x q^{(n-x)}$$

is straightforward, it can be tedious, especially when n and x are large and when there are a large number of terms in 
the summation. It is not surprising that a great deal of attention and ingenuity have been applied to constructing 
useful approximations for sums of this kind. 

## Applications 

1. The binomial distribution arises whenever underlying events have two possible outcomes, the chances of which remain
constant. The importance of the distribution has extended from its original application in gaming to many other areas. 

2. Its use in genetics arises because the inheritance of biological characteristics depends on genes that occur in pairs;
see, for example, Fisher and Matheras (1936) analysis of data on straight versus wavy hair in mice.

3. More recent application in genetics is the study of the number of nucleotides that are in the same state in two 
DNA sequences (Kaplan and Risko, 1982). 

4. The number of defectives found in random samples of size n from a stable production process is a binomial
variable; acceptance sampling is a very important application of the test for the mean of a binomial sample against a
hypothetical value. 

5. Seber (1982b) has given a number of instances of the use of the binomial distribution in animal ecology, for example, 
in mark-recapture estimation of the size of an animal population. Boswell, Ord, and Patil (1979) gave applications in 
plant ecology. 

Although appealing in their simplicity, the assumptions of independence and constant probability for the binomial
distribution are not often precisely satisfied. Published critical appraisals of the extent of departure from these
assumptions in actual situations are rather rare.

# Binomial Mixture Distributions

The binomial distribution has two parameters, $n$ and $p$, and either or both of these may be supposed to have a probability
distribution. We will not discuss cases in which both $n$ and $p$ vary, though it is easy to construct such examples. 

In most cases discussed in the statistical literature, $p$ has a continuous distribution, while $n$ is discrete. The 
latter restriction is necessary, but the former is not.

However, discrete distributions for $p$ have not been found to be useful and have not attracted much attention from a
theoretical point of view. 

# Main Groups for Binomial Distribution

More than 50 distributions are in this blog post but I shall not explain 
them even briefly. This blog post is to introduce them and let you know the 
reader about these distributions with article references.

In this section I introduce the main groups, where the number of groups are 6. Further,
the Binomial Parent group has a very broad approach. "Binomial Parent" means, 
Where Binomial Distribution is the Parent distribution and by mixing them with other distributions
we can produce a new distribution which would satisfy real data but not fully satisfy the condition 
of Binomial Distribution.

Clearly with the vast number of distributions and sub groups it is not possible to draw them in one
diagram, therefore I have divided them in sub groups and sub topics in the following sections.

```{r Diagramme,echo=FALSE}
library(DiagrammeR)

DiagrammeR(
           "graph TB;
            A[Binomial <br> Distribution]-->B[Binomial <br> Not Parent];
            A-->D[Other with <br> no sub groups];
            A-->C[Binomial <br> Parent];
            A-->M[Alternate Binomial <br> Distribution];
            A-->N[Neyman Type A <br> Distribution];
            A-->O[Hermite Distribution];
            C-->CA[Mixing Parameter];
            CA-->CAA[N/n <br> Binomial];
            CA-->CAB[K <br> Binomial];
            CA-->CAC[Y <br> Binomial];
            CA-->CAE[p transformed <br> Binomial];
            CA-->CAF[Cumulative Distribution <br> Function];
            CA-->CAD[p <br> Binomial];
            CA-->CAG[Log Inverse <br> Distribution <br> 0,1 Domain]
           ",width=1000,height=400)
```

## Other With No Sub-Groups

First sub group in discussion is "Other With No Sub-Groups", which is an abbreviation for Binomial distributions
which were direcly created for the purpose of satisfying specific real world situations and theoretical concepts.

```{r Other with No Sub-Groups,echo=FALSE}
DiagrammeR(
           "graph TB;
            B[Binomial <br> Distribution]-->A[Other with <br> no sub groups];
            A-->D[Grassia Binomial <br> Distribution];
            A-->E[Zero Modified <br> Binomial Distribution];
            A-->F[Dandekar's Modified <br> Binomial Distribution];
            A-->G[Simplex Binomial <br> Mixture Model];
            A-->H[Double Binomial <br> Distribution];
            A-->I[Finite Binomial <br> Mixtures];
            A-->J[Binomial Distribution <br> of order K];
            A-->K[Truncated Binomial <br> Distribution];
            A-->L[Weighted Binomial <br> Distribution];
            ",width=1000,height=400)
```

Below is the list of Distributions in this sub group with article references.

* Grassia Binomial Distribution

* Zero Modified Binomial Distribution

* Dandekar's Modified Binomial Distribution

* Simplex Binomial Mixture Model

* Double Binomial Distribution

* Finite Biomial Mixtures

* Binomial Distribution of order K

* Truncated Binomial Distribution

* Weighted Binomial Distribution

## Binomial Not Parent

This is the sub group where Binomial Distribution is not the parent but
rather some other different distribution like Hypergeometric or Negative Binomial or Poisson.

```{r Binomial not parent,echo=FALSE}
DiagrammeR(
           "
            graph TB;
            A[Binomial <br> Distribution]-->B[Binomial Not Parent];
            B-->BA[Hypergeometric <br> + Binomial];
            B-->BB[Negative Binomial <br> + Binomial];
            B-->BC[Poisson + <br> Binomial];
           ",width=800,height=400
)
```

* Hypergeometric + Binomial

* Negative Binomial + Binomial

* Poisson + Binomial

## Alternate Binomial Distributions

This is a sub group of distributions which can be used as an alternative to
Binomial Distribution.

```{r Alternate Binomial Distributions,echo=FALSE}
DiagrammeR(
           "
            graph TB;
            A[Binomial <br> Distribution]-->B[Alternate Binomial <br> Distribution];
            B-->C[Additive Binomial <br> Distribution];
            B-->D[Beta-Correlated <br> Binomial Distribution];
            B-->E[COM-Poisson Binomial <br> Distribution];
            B-->F[Correlated Binomial <br> Distribution];
            B-->G[Multiplicative Binomial <br> Distribution];
            ",width=800,height=400
)
```

* Additive Binomial Distribution

* Beta-Correlated Binomial Distribution

* COM-Poisson Binomial Distribution

* Correlated Binomial Distribution

* Multiplicative Binomial Distribution

## Neyman Type A Distribution 

```{r Neyman Type A Distribution,echo=FALSE}
DiagrammeR(
           "
            graph TB;
            A[Binomial <br> Distribution]-->B[ Neyman Type A <br> Distribution];
            B-->C[Poisson + Binomial + Binomial + Beta];
            B-->D[Poisson + Poisson + Binomial + Beta <br> or Poisson + Binomial + Poisson + Beta];
            B-->E[Poisson + Binomial + Poisson + Beta <br> or Binomial + Poisson + Poisson + Beta];
           ",width=800,height=400
)
```

* Poisson + Binomial + Binomial + Beta

* Poisson + Poisson + Binomial + Beta <br> or Poisson + Binomial + Poisson + Beta

* Poisson + Binomial + Poisson + Beta <br> or Binomial + Poisson + Poisson + Beta

## Hermite Distribution

```{r Hermite Distribution,echo=FALSE}
DiagrammeR(
           "
           graph TB;
           A[Binomial <br> Distribution]-->B[Hermite <br> Distribution];
           B-->C[Binomial + Poisson];
           B-->D[Poisson + Binomial];
           ",width=800,height=400
)
```

* Binomial + Poisson

* Poisson + Binomial


## Binomial Parent 


### N/n , K and Y Mixtures

```{r other Mixtures,echo=FALSE}
DiagrammeR(
           "
           graph TB;
           A[Binomial <br> Distribution]-->B[Binomial <br> Parent];
           B-->C[Mixing Parameter];
           C-->D[N/n <br> Binomial];
           D-->DA[Poisson];
           D-->DB[Binomial];
           D-->DC[Negative Binomial];
           D-->DD[Logarithmic];
           C-->E[K <br> Binomial];
           E-->EA[Poisson];
           C-->F[Y <br> Binomial];
           F-->FA[Hypergeometric];
           ",width=800,height=400
)
```


### p Transformed Binomial

```{r Transformed Binomial,echo=FALSE}
DiagrammeR(
           "
           graph LR;
           A[Binomial <br> Distribution]-->B[Binomial <br> Parent];
           B-->C[Mixing Parameter];
           C-->D[p transformed <br> Binomial];
           D-->E[p = 1 - exp_-t_ ];
           E-->EA[Binomial Exponential <br> Distribution];
           E-->EB[Binomial Gamma 1 <br> Distribution];
           E-->EC[Binomial Gamma 2 <br> Distribution];
           E-->ED[Binomial Generalized <br> Exponential 1 Distribution];
           E-->EE[Binomial Generalized <br> Exponential 2 Distribution];
           D-->F[p = exp_-t_ ];
           F-->FA[Binomial Exponential <br> Distribution];
           F-->FB[Binomial Gamma 1 <br> Distribution];
           F-->FC[Binomial Gamma 2 <br> Distribution];
           F-->FD[Binomial Generalized <br> Exponential 1 Distribution];
           F-->FE[Binomial Generalized <br> Exponential 2 Distribution];
           F-->FF[Binomial Variated <br> Exponential Distribution];
           F-->FG[Binomial Variated <br> Gamma 2,alpha Distribution];
           F-->FH[Binomial Inverse <br> Gaussian Distribution];
           D-->G[p = cy];
           G-->GA[Binomial Generalized Beta 4 Distribution];
           ",width=800,height=1000
)
```


### Log Inverse Distribution [0,1] Domain

```{r Log Inverse Distribution 0,1 Domain,echo=FALSE}
DiagrammeR(
           "
           graph LR;
           A[Binomial <br> Distribution]-->B[Binomial <br> Parent];
           B-->C[Mixing Parameter];
           C-->D[Log Inverse <br> Distribution <br> 0,1 Domain];
           D-->DA[Binomial Type 1 <br> Log Inverse <br> Exponential Distribution];
           D-->DB[Binomial Type 2 <br> Log Inverse <br> Exponential Distribution];
           D-->DC[Binomial Type 1 <br> Log Inverse <br> Gamma 1 Distribution];
           D-->DD[Binomial Type 2 <br> Log Inverse <br> Gamma 1 Distribution];
           D-->DE[Binomial Type 1 <br> Log Inverse <br> Gamma 2 Distribution];
           D-->DF[Binomial Type 2 <br> Log Inverse <br> Gamma 2 Distribution];
           D-->DG[Binomial Type 1 <br> Log Inverse <br> Generalized <br> Exponential 1 <br> Distribution];
           D-->DH[Binomial Type 2 <br> Log Inverse <br> Generalized <br> Exponential 1 <br> Distribution];
           D-->DJ[Binomial Type 1 <br> Log Inverse <br> Generalized <br> Exponential 2 <br> Distribution];
           D-->DK[Binomial Type 2 <br> Log Inverse <br> Generalized <br> Exponential 2 <br> Distribution];
           D-->DL[Binomial Type 1 <br> Log Inverse <br> Gamma 3 <br> Distribution];
           ",width=800,height=1000
)
```


### Cumulative Distribution Function

```{r Cumulative Distribution Function,echo=FALSE}
DiagrammeR(
           "
            graph TB;
            A[Binomial <br> Distribution]-->B[Binomial Parent];
            B-->C[Mixing Parameter];
            C-->D[Cumulative Distribution];
            D-->DA[Beta Generated <br> Distribution];
            DA-->DAA[Binomial <br> Beta Exponential <br> Distribution];
            DA-->DAB[Binomial <br> Beta Generalized <br> Exponential Distribution];
            DA-->DAC[Binomial <br> Beta Power <br> Distribution];
            D-->DB[Kumaraswamy <br> Generated <br> Distribution];
            DB-->DBA[Binomial <br> Kumaraswamy <br> Power Distribution];
            DB-->DBB[Binomial <br> Kumaraswamy <br> Exponential <br> Distribution];        
           ",width=1000,height=750
)
```

### p Binomial 

```{r p Binomial,echo=FALSE}
DiagrammeR(
           "
           graph LR;
           A[Binomial <br> Distribution]-->B[ Binomial <br> Parent];
           B-->C[Mixing Parameter];
           C-->D[p <br> Binomial];
           D-->DA[Beyond Beta <br> Distribution];
           DA-->DAA[Binomial Triangular <br> Distribution];
           DA-->DAB[Binomial Kumaraswamy 2 <br> Distribution];
           DA-->DAC[Binomial Kumaraswamy 1 <br> Distribution];
           DA-->DAD[Binomial Truncated <br> Exponential Distribution];
           DA-->DAE[Binomial Truncated <br> Gamma Distribution];
           DA-->DAF[Binomial - MinusLog <br> Distribution];
           DA-->DAG[Binomial Standard <br> Two Sied Power <br> Distribution];
           DA-->DAH[Binomial Ogive <br> Distribution];
           DA-->DAI[Binomial - Two Sided <br> Ogive Distribution];
           D-->DB[Beta Distribution];
           DB-->DBA[Beta - Binomial <br> Distribution];
           DB-->DBB[McDonald Generalized <br> Beta - Binomial <br> Distribution];
           DB-->DBC[Libby and Novick <br> Generalized Beta - Binomial <br> Distribution];
           DB-->DBD[Gauss Hypergeometric <br> Binomial Distribution];
           DB-->DBE[Confluent Hypergeometric <br> Binomial Distribution];
           DB-->DBF[Binomial Uniform <br> Distribution];
           DB-->DBG[Binomial Power <br> Function Distribution];
           DB-->DBH[Binomial Truncated <br> Beta Distribution];
           DB-->DBI[Binomial Arcsine <br> Distribution];
           ",width=700,height=1200
)
```

# References 