---
title: Benchmarking maxLik function
author: M.Amalan
date: '2018-12-20'
slug: benchmarking-maxlik-function
categories:
  - fitODBOD
tags:
  - Beta-Binomial
  - fitODBOD
  - maxLik
image:
  caption: ''
  focal_point: ''
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE,comment = NA,fig.height = 10,fig.width = 10,warning = FALSE,message = FALSE)
library(tidyverse)
library(formattable)
library(kableExtra)
```

# Estimating the shape parameters of Beta-Binomial Distribution

# Introduction

Beginning of this month I wrote a 
[small section](https://amalan-con-stat.netlify.com/post/fitodbod_1/benchmarking-maximum-liklihood-functions-from-r/#maxlik-function) 
regarding maxLik function by comparing it to other optimization functions. Here, we will further study the analytical methods
which can be used in this function and compare them to find suitability. maxLik function is from the package 
[maxLik](https://cran.r-project.org/web/packages/maxLik/index.html). Further, 
[Documentation](https://www.rdocumentation.org/packages/maxLik/versions/1.3-4/topics/maxLik)
clearly indicates all things related to the function in detail. 

Focusing on the seven analytical methods is my intention from this blog post. So, we have the Beta-Binomial distribution and Binomial 
Outcome data, and need to estimate proper shape parameters which would Maximize the Log Likelihood value of the Beta-Binomial 
distribution for the above Binomial Outcome data. In this case Alcohol Consumption data from the 
[fitODBOD](https://cran.r-project.org/package=fitODBOD) package will be used.

Further we will focus on the process time to optimization, estimated shape parameters, maximized Log Likelihood value, 
expected frequencies, p-value and Over-dispersion with tables.

Below are the seven analytical methods in concern

1. NR
2. BFGS
3. BFGSR
4. BHHH
5. SANN
6. CG
7. NM

Alcohol Consumption data has two sets of frequency values but only values from week 1 will be used. Below is the the Alcohol 
Consumption data, where number of observations is 399 and the Binomial Random variable is a vector of values from zero to seven.

```{r Data}
library(fitODBOD)
kable(Alcohol_data,"html",align=c('c','c','c')) %>%
  kable_styling(bootstrap_options = c("striped"),font_size = 14,full_width = F) %>%
  row_spec(0,color = "blue") %>%
  column_spec(1,color = "red")
```

## Brief of maxLik Function

Small [section](https://amalan-con-stat.netlify.com/post/fitodbod_1/benchmarking-maximum-liklihood-functions-from-r/#maxlik-function)
about the maxLik function will be very useful to understand this blog post. 

Reference : Henningsen, A. and Toomet, O. (2011): maxLik: A package for maximum likelihood estimation in R Computational Statistics 26, 
443--458 Marquardt, D.W., (1963) An Algorithm for Least-Squares Estimation of Nonlinear Parameters, 
Journal of the Society for Industrial & Applied Mathematics 11, 2, 431--441

So for the initial parameters of a=0.1 and b=0.2 we will be finding estimated parameters from different analytical methods which 
would maximize the Log Likelihood value of the Beta-Binomial distribution.

Here also we are transforming the given EstMLEBetaBin function to satisfy the maxLik function conditions.

```{r maxLik function from stats package}
# new function to facilitate maxLik criteria
# only one input but has two elements
formaxLik<-function(a)
  {
  -EstMLEBetaBin(x=Alcohol_data$Days, freq=Alcohol_data$week1,a=a[1],b=a[2])
  }
```

So the formaxLik function can be used as above and parameters are estimated for $\alpha$ and $\beta$ (or a, b) for the Alcohol 
Consumption data week 1. Further the maxLik function can be scrutinized as below.

* package : maxLik 
* No of Inputs: 6
* Minimum required Inputs : 2
* Class of output : list or class of maxim or class of maxLik
* No of outputs: 11
* No of Analytical Methods : 7
* Default Method : Automatically chosen

## NR method

NR is an abbreviation for Unconstrained and equality-constrained maximization based on the quadratic approximation (Newton) method.
The idea of the Newton method is to approximate the function at a given location by a multidimensional quadratic function, and use 
the estimated maximum as the start value for the next iteration.

```{r maxLik function NR method, results='hide'}
library(maxLik)

# optimizing values for a,b using NR analytical method
NR_answer<-maxLik(logLik =formaxLik, start = c(0.1,0.2),method = "NR")

# obtaining class of output
class(NR_answer)

# length of output
length(NR_answer)

# the outputs
NR_answer$estimate # estimated values for a, b
NR_answer$maximum # minimized function value 
NR_answer$iterations  # no of iterations to succeed
NR_answer$gradient # last gradient value which was calculated
NR_answer$message # additional information
NR_answer$hessian # hessian matrix
NR_answer$code # indicates successful completion
NR_answer$fixed # logical vector indicating which parameters are constants
NR_answer$type # type of maximization
NR_answer$last.step # list describing the last unsuccessful step
NR_answer$control # see the documentation understand

# fitting the Beta-Binomial distribution with estimated shape parameter values
fitBetaBin(Alcohol_data$Days,Alcohol_data$week1,
           NR_answer$estimate[1],NR_answer$estimate[2])
```

## BFGS method

BFGS is a Quasi-Newton method (also known as a variable metric algorithm), specifically that published simultaneously in 1970 by Broyden, 
Fletcher, Goldfarb and Shanno. This uses function values and gradients to build up a picture of the surface to be optimized.

Reference : Broyden, C.G., 1967. Quasi-Newton methods and their application to function minimization. Mathematics of Computation, 21(99), pp.368-381.

```{r maxLik function BFGS method, results='hide'}
# optimizing values for a,b using BFGS analytical method
BFGS_answer<-maxLik(logLik =formaxLik, start = c(0.1,0.2),method = "BFGS")

# obtaining class of output
class(BFGS_answer)

# length of output
length(BFGS_answer)

# the outputs
BFGS_answer$estimate # estimated values for a, b
BFGS_answer$maximum # minimized function value 
BFGS_answer$iterations  # no of iterations to succeed
BFGS_answer$gradient # last gradient value which was calculated
BFGS_answer$message # additional information
BFGS_answer$hessian # hessian matrix
BFGS_answer$code # indicates successful completion
BFGS_answer$fixed # logical vector indicating which parameters are constants
BFGS_answer$type # type of maximization
BFGS_answer$last.step # list describing the last unsuccessful step
BFGS_answer$control # see the documentation understand

# fitting the Beta-Binomial distribution with estimated shape parameter values
fitBetaBin(Alcohol_data$Days,Alcohol_data$week1,
           BFGS_answer$estimate[1],BFGS_answer$estimate[2])
```


## BFGSR method

```{r maxLik function BFGSR method, results='hide'}
# optimizing values for a,b using BFGSR analytical method
BFGSR_answer<-maxLik(logLik =formaxLik, start = c(0.1,0.2),method = "BFGSR")

# obtaining class of output
class(BFGSR_answer)

# length of output
length(BFGSR_answer)

# the outputs
BFGSR_answer$estimate # estimated values for a, b
BFGSR_answer$maximum # minimized function value 
BFGSR_answer$iterations  # no of iterations to succeed
BFGSR_answer$gradient # last gradient value which was calculated
BFGSR_answer$message # additional information
BFGSR_answer$hessian # hessian matrix
BFGSR_answer$code # indicates successful completion
BFGSR_answer$fixed # logical vector indicating which parameters are constants
BFGSR_answer$type # type of maximization
BFGSR_answer$last.step # list describing the last unsuccessful step
BFGSR_answer$control # see the documentation understand

# fitting the Beta-Binomial distribution with estimated shape parameter values
fitBetaBin(Alcohol_data$Days,Alcohol_data$week1,
           BFGSR_answer$estimate[1],BFGSR_answer$estimate[2])
```

## BHHH method



## SANN method

```{r maxLik function SANN method, results='hide'}
# optimizing values for a,b using SANN analytical method
SANN_answer<-maxLik(logLik =formaxLik, start = c(0.1,0.2),method = "SANN")

# obtaining class of output
class(SANN_answer)

# length of output
length(SANN_answer)

# the outputs
SANN_answer$estimate # estimated values for a, b
SANN_answer$maximum # minimized function value 
SANN_answer$iterations  # no of iterations to succeed
SANN_answer$gradient # last gradient value which was calculated
SANN_answer$message # additional information
SANN_answer$hessian # hessian matrix
SANN_answer$code # indicates successful completion
SANN_answer$fixed # logical vector indicating which parameters are constants
SANN_answer$type # type of maximization
SANN_answer$last.step # list describing the last unsuccessful step
SANN_answer$control # see the documentation understand

# fitting the Beta-Binomial distribution with estimated shape parameter values
fitBetaBin(Alcohol_data$Days,Alcohol_data$week1,
           SANN_answer$estimate[1],SANN_answer$estimate[2])
```

## CG method

According to the documentation Method CG is a conjugate gradients method based on that by Fletcher and Reeves (1964) 
(but with the option of Polak-Ribiere or Beale-Sorenson updates). Conjugate gradient methods will generally be more fragile 
than the BFGS method, but as they do not store a matrix they may be successful in much larger optimization problems.

Reference : Fletcher, R. and Reeves, C.M., 1964. Function minimization by conjugate gradients. The computer journal, 7(2), pp.149-154.

```{r maxLik function CG method, results='hide'}
# optimizing values for a,b using CG analytical method
CG_answer<-maxLik(logLik =formaxLik, start = c(0.1,0.2),method = "CG")

# obtaining class of output
class(CG_answer)

# length of output
length(CG_answer)

# the outputs
CG_answer$estimate # estimated values for a, b
CG_answer$maximum # minimized function value 
CG_answer$iterations  # no of iterations to succeed
CG_answer$gradient # last gradient value which was calculated
CG_answer$message # additional information
CG_answer$hessian # hessian matrix
CG_answer$code # indicates successful completion
CG_answer$fixed # logical vector indicating which parameters are constants
CG_answer$type # type of maximization
CG_answer$last.step # list describing the last unsuccessful step
CG_answer$control # see the documentation understand

# fitting the Beta-Binomial distribution with estimated shape parameter values
fitBetaBin(Alcohol_data$Days,Alcohol_data$week1,
           CG_answer$estimate[1],CG_answer$estimate[2])
```

## NM method

NM is the abbreviation to Nelder and Mead method. According to the documentation it uses only function values and 
is robust but relatively slow. It will work reasonably well for non-differential functions.

Reference : Nelder, J.A. and Mead, R., 1965. A simplex method for function minimization. The computer journal, 7(4), pp.308-313.

```{r maxLik function NM method, results='hide'}
# optimizing values for a,b using NM analytical method
NM_answer<-maxLik(logLik =formaxLik, start = c(0.1,0.2),method = "NM")

# obtaining class of output
class(NM_answer)

# length of output
length(NM_answer)

# the outputs
NM_answer$estimate # estimated values for a, b
NM_answer$maximum # minimized function value 
NM_answer$iterations  # no of iterations to succeed
NM_answer$gradient # last gradient value which was calculated
NM_answer$message # additional information
NM_answer$hessian # hessian matrix
NM_answer$code # indicates successful completion
NM_answer$fixed # logical vector indicating which parameters are constants
NM_answer$type # type of maximization
NM_answer$last.step # list describing the last unsuccessful step
NM_answer$control # see the documentation understand

# fitting the Beta-Binomial distribution with estimated shape parameter values
fitBetaBin(Alcohol_data$Days,Alcohol_data$week1,
           NM_answer$estimate[1],NM_answer$estimate[2])
```


# Sumary of Time evalutation for different Analytical methods of maxLik function

# Summary of results after using the maxLik function for different analytical methods 

# Final Conclusion

_Thank you_ 