---
title: 'Sri Lanka and its affect on/in Journal Articles '
author: M.Amalan
date: '2019-02-18'
slug: SLandJournal
output:
    blogdown::html_page:
      toc : true
categories:
  - Blog
  - R
  - SL
  - Journal
tags:
  - gganimate
  - tidyverse
  - R package
  - R
  - SL
image:
  caption: ''
  focal_point: ''
summary: "Exploring How has Sri Lanka done in related to Academic Journals."
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE,message = FALSE,warning = FALSE,fig.width = 12,fig.height = 12)
```

# Introduction

"SCImago Journal and Country Rank provides valuable estimates of academic journals' presitige.", this is 
according to its GitHub repository which could be found [here.](https://github.com/ikashnitsky/sjrdata)
I completely agree with this statement, last year(2018) while searching for journals to publish a research
of my interest and contribution this [website](https://www.scimagojr.com/) came in useful comparative to
most of my google searches.

It is crucial to find a journal which summarizes the research work we intend to contribute with more
relative information such as ranking, h-index, citations and etc which could be compared for our benefit.
Usually people tend to collect information by themselves and might miss a valuable and more suitable
opportunity to publish because they did not know where to find them. That will not occur if you use 
SciMagoJr.

# Packages and Ideas

As a Researcher from Sri Lanka I am much more interested in how we as a country has made progress
in related to journals or research publishing in perspective of our neighboring countries and fields
of research interest.

The package sjrdata has three data-sets and they provide information regarding to journals
which is mostly seeked by researchers to find an appropriate place to publish their research
or find places which would benefit them in finding a unique research problem. 

```{r Packages and data}
# Data
library(sjrdata)

# Packages
library(tidyverse)
library(gganimate)
library(ggrepel)
library(magrittr)
library(lubridate)
library(splitstackshape)
library(qdap)
library(ggthemr)
ggthemr("flat dark")
```

The three datasets in concern are 

1.  sjr_countries - Contains information regarding h_index over the years for countries and their ranks 
                    with other information related to documents.
                    
2.  sjr_journals - Records for type of document with their basic information which can be connected to country of
                   origin, publisher and fields of interest and much more.
                   
3.  sjr_countries_1996_2017 - Similar to sjr_countries but this is an accumulation of the data-set which would 
                              indicate the overall perfromace as a country from 1996 to 2017.

# Country Information

This section will simply scrutinize how Sri Lanka over the years has performed and it should be noted that only 
the ranks of South Asian Region countries will be compared for the purpose of understanding the progress 
as a whole community. 

## Ranking of Sri Lanka from 1996 to 2017

Sri Lanka(SL) had its ups and downs in the rankings where the worst place was to be in
85th and the best place to be in 75th. As of in 2017 it has reached 78th which is good
considering the previous three years(2014-2016) the ranks were higher.

In 1996 SL was holding 85th and this occurs again in the years 2011 and 2014. Also it 
reached the 75th ranking position in year 2007. I sincerely hope that in the coming years 
SL would reach better ranks.

```{r sjr Sri Lanka and rank}
subset(sjr_countries,country=="Sri Lanka") %>%
  mutate(year=year(as.Date(year,"%Y"))) %>%
ggplot(.,aes(x=year,y=rank,label=rank))+
  geom_point()+geom_line()+
  geom_text_repel()+
  xlab("Year")+ylab("Rank")+
  ggtitle("Rank of Sri Lanka Changing from 1996 to 2017")+
  theme(axis.text.x =element_text(angle = 45, hjust = 1))
```

## Ranking of South Asian Region Countries from 1996 to 2017



```{r sjr South Asia and rank}
subset(sjr_countries,
       country =="Sri Lanka" | country=="India"| country=="Pakistan"|
       country =="Nepal" | country== "Maldives" | country =="Afghanistan"|
       country == "Bhutan" | country=="Bangladesh") %>%
       mutate(year=year(as.Date(year,"%Y"))) %>%
ggplot(.,aes(x=year,y=rank,label=rank,color=country))+
  geom_point()+geom_line()+
  geom_text_repel()+
  xlab("Year")+ylab("Rank")+
  ggtitle("Rank of South Asian Region Changing from 1996 to 2017")+
  theme(axis.text.x =element_text(angle = 45, hjust = 1))
```


## Documents , Citable Documents and Self Citations from Sri Lanka Journals 

```{r sjr Sri Lanka and documents citable documents self citations}
sjr_countries%>%
  gather(citing,values,"documents","citable_documents","self_citations") %>%
  subset(country=="Sri Lanka",select=c(citing,values,year)) %>%
  mutate(year=year(as.Date(year,"%Y"))) %>%
ggplot(.,aes(x=year,y=values,label=values,color=citing,shape=citing))+
  geom_point()+geom_line()+
  geom_text_repel()+
  xlab("Year")+ylab("Counts")+
  scale_x_continuous(breaks=seq(1996,2017),labels=seq(1996,2017))+
  scale_y_continuous(breaks=seq(0,2000,100),labels=seq(0,2000,100))+
  ggtitle("Document Counts Changing from 1996 to 2017 for Sri Lanka")+
  theme(axis.text.x =element_text(angle = 90, hjust = 1))
```

## Citations per Document from Sri Lankan Journals

```{r sjr citations per document}
sjr_countries %>%
   subset(country=="Sri Lanka") %>%
   mutate(year=year(as.Date(year,"%Y"))) %>%
ggplot(.,aes(x=year,y=citations_per_document,
          label=citations_per_document))+
  geom_point()+geom_line()+
  geom_text_repel()+
  xlab("Year")+ylab("Citations Per Document")+
  scale_x_continuous(breaks=seq(1996,2017),labels=seq(1996,2017))+
  scale_y_continuous(breaks=seq(0,30),labels=seq(0,30))+
  ggtitle("Citations Per Document from 1996 to 2017 for Sri Lanka")+
  theme(axis.text.x =element_text(angle = 90, hjust = 1))
```

# Journals from Sri Lanka

## Rank changes for Sri Lankan Journals

```{r rank changes}
p<-subset(sjr_journals,country=="Sri Lanka") %>%
   mutate(year=year(as.Date(year,"%Y"))) %>%
   ggplot(.,aes(x=str_wrap(factor(title),30),y=rank,color=year))+
         geom_jitter()+coord_flip()+
         shadow_mark()+
         transition_time(year)+ease_aes("linear")+
         xlab("Title of Journals")+ylab("Rank")+
         ggtitle("Ranks Changing over time for Sri Lankan Journals from 1999 to 2017",
                 subtitle = "Year : {round(frame_time)}")

animate(p,fps=1,nframes=18)  
```

## SJR values changing for Sri Lankan Journals

```{r sjr changes}
p<-subset(sjr_journals,country=="Sri Lanka") %>%
   mutate(year=year(as.Date(year,"%Y"))) %>%
   ggplot(.,aes(x=str_wrap(factor(title),30),y=sjr,color=year,shape=sjr_best_quartile))+
         geom_jitter()+coord_flip()+
         shadow_mark()+
         transition_time(year)+ease_aes("linear")+
         xlab("Title of Journals")+ylab("SJR values")+
         ggtitle("SJR values Changing over time for Sri Lankan Journals from 1999 to 2017",
                 subtitle = "Year : {round(frame_time)}")+
         theme(legend.position = "bottom")

animate(p,fps=1,nframes=18)  
```

## References per Document changing for Sri Lankan Journals

```{r ref per doc changes}
p<-subset(sjr_journals,country=="Sri Lanka") %>%
   mutate(year=year(as.Date(year,"%Y"))) %>%
   ggplot(.,aes(x=str_wrap(factor(title),30),y=ref_doc,color=year))+
         geom_jitter()+coord_flip()+
         shadow_mark()+
         transition_time(year)+ease_aes("linear")+
         xlab("Title of Journals")+ylab("References per Document")+
         ggtitle("References per Document Changing over time for Sri Lankan Journals from 1999 to 2017",
                 subtitle = "Year : {round(frame_time)}")

animate(p,fps=1,nframes=18)  
```

## Publishers and Journals from Sri Lanka
 
```{r publisher and Journals}
p<-subset(sjr_journals,country=="Sri Lanka") %>%
   mutate(year=year(as.Date(year,"%Y"))) %>%
   ggplot(.,aes(x=str_wrap(factor(title),30),y=str_wrap(factor(publisher),30),color=year))+
   geom_jitter()+
   transition_time(year)+ease_aes("linear")+
   shadow_mark()+
   xlab("Title")+ylab("Publisher")+
   theme(axis.text.x =element_text(angle = 90, hjust = 1))+
   ggtitle("Title and Publishers changing over time for Sri Lankan Journals from 1999 to 2017",
           subtitle = "Year : {round(frame_time)}")

animate(p,fps=1,nframes=18) 
```
 
## Total Documents and References for Journals from Sri Lanka

```{r tot doc and ref}
p<-sjr_journals %>%
    gather(Group,values,"total_docs_year","total_refs") %>%
    subset(country=="Sri Lanka",select=c(Group,values,year,title)) %>%
    mutate(year=year(as.Date(year,"%Y"))) %>%
ggplot(.,aes(x=str_wrap(factor(title),30),y=values,shape=Group,color=year))+
      geom_jitter()+coord_flip()+
      xlab("Title")+ylab("Count")+
      transition_time(year)+ease_aes("linear")+
      shadow_mark()+
      ggtitle("Counts changing for Sri Lankan Journals from 1999 to 2017",
              subtitle = "Year : {round(frame_time)}")+
      theme(legend.position="bottom")

animate(p,nframes=18,fps=1)
```

## Three year time period for Journals from Sri Lanka

```{r 3 doc cite citable year}
p<-sjr_journals %>%
    gather(Group,values,"total_docs_3years","total_cites_3years","citable_docs_3years") %>%
    subset(country=="Sri Lanka",select=c(Group,values,year,title)) %>%
    mutate(year=year(as.Date(year,"%Y"))) %>%
ggplot(.,aes(x=str_wrap(factor(title),30),y=values,shape=Group,color=year))+
      geom_jitter()+coord_flip()+
      xlab("Title")+ylab("Count for 3 years")+
      transition_time(year)+ease_aes("linear")+
      shadow_mark()+
      ggtitle("Counts changing for Sri Lankan Journals from 1999 to 2017",
              subtitle = "Year : {round(frame_time)}")+
      theme(legend.position="bottom")

animate(p,nframes=18,fps=1)
```

## Journals and categories


```{r journals and categories}
p<-sjr_journals %>%
      subset(country=="Sri Lanka") %>%
      select(year,title,categories) %>%
      cSplit("categories",sep=";") %>%
      gather(Groups,Categories,"categories_1","categories_2","categories_3","categories_4") %>%
      mutate(Categories=genX(Categories,"(Q",")")) %>%
      mutate(year=year(as.Date(year,"%Y"))) %>%
ggplot(.,aes(x=str_wrap(factor(title),30),y=str_wrap(Categories,30),shape=Groups,color=year)) +
      geom_jitter()+ 
      theme(axis.text.x =element_text(angle = 90, hjust = 1),
            legend.position="bottom")+
      transition_time(year)+ease_aes("linear")+
      xlab("Title")+ ylab("Categories")+
      ggtitle("Title and Categories between the years 1999 to 2017",
              subtitle="Year :{round(frame_time)}")

animate(p,nframes=18,fps=1)             
```

*THANK YOU*