---
title: 'Week 9 : French Train Delays'
author: M.Amalan
date: '2019-02-28'
slug: Week9
ouptut:
  blogdown::html_page:
    toc: true
categories:
  - TidyTuesday
tags:
  - tidyverse
  - TidyTuesday
  - gganimate
image:
  caption: ''
  focal_point: ''
summary: "2019 Week 9 TidyTuesday: Train Delays in French Land"  
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE,warning = FALSE, message = FALSE,fig.width = 12,fig.height = 12)
```

```{r load the data and packages}
# load the packages
library(readr)
library(tidyverse)
library(gganimate)
library(ggalluvial)
library(geomnet)
library(ggthemr)

# load the theme
ggthemr("fresh")

# load the data
small_trains <- read_csv("small_trains.csv")
```

# Network Graph for the French City Trains

Simply drawing a network graph to understand which french cities are mainly urban with 
capacity to trains arriving and departuring. Cities such as Paris Lyon, Paris Montparnasse,
Paris Nord and Paris Est could be cities of concern with much for traffic with related to trains.

```{r network graph, fig.width= 14,fig.height=14}
ggplot(small_trains,aes(from_id=departure_station,to_id=arrival_station))+
          geom_net(directed = TRUE,labelon = TRUE,size=0.5,labelcolour = "black",
                   repel = FALSE,ecolour = "grey70", arrowsize = 0.75,
                   linewidth = 0.5,layout.alg = "fruchtermanreingold")+
          theme_net()+
          ggtitle("Network Graph Showing from City to City of French Trains")

```

# Paris Montparnasse

Let me focus on Montparnasse which has lot of trains coming towards and departuring outwards
according to the network map. Not all are cities of France according to my observations, where
I can see Madrid, Zurich and Barcelona. 

# Chosen City with Total Number of Trips

Total number of trips from Paris Montparnasse station to other cities is noted here.
Four years of data with more accuracy by months is considered in this plot.
There is clear variation in this data for cities. 

Close to 800 trips have been recorded towards the Bordeaux St Jean city but not clearly the same
pattern for all years or months as well. Further, St Malo city has the lowest amount of trips
compared to other cities in all fours but follows a centered pattern around the 100 mark.

```{r Paris Montparnasse total num of trips}
p<-ggplot(subset(small_trains,departure_station=="PARIS MONTPARNASSE"),
       aes(x=str_wrap(arrival_station,20),y=total_num_trips,color=month))+
       geom_jitter()+coord_flip()+ labs(color="Month")+
       transition_time(year)+ease_aes("linear")+
       scale_y_continuous(breaks = seq(0,800,100),labels=seq(0,800,100))+
       xlab("Arrival Station")+ylab("Total Number of Trips")+
       ggtitle("Paris Montparnasse and its arrival Stations" ,subtitle ="Year :{frame_time}")

animate(p,nframes=4,fps=1)
```

# Chosen City with Average Journey Time

```{r Paris Montparnasse Journey time average}
p<-ggplot(subset(small_trains,departure_station=="PARIS MONTPARNASSE"),
       aes(x=str_wrap(arrival_station,20),y=journey_time_avg,color=month))+
       geom_jitter()+coord_flip()+ labs(color="Month")+
       transition_time(year)+ease_aes("linear")+
       xlab("Arrival Station")+ylab("Average Journey Time")+
       ggtitle("Paris Montparnasse and its arrival Station" ,subtitle ="Year :{frame_time}")

animate(p,nframes=4,fps=1)
```

# Chosen City with Average Delay with All Departing

```{r Paris Montparnasse avg delay all departing}
p<-ggplot(subset(small_trains,departure_station=="PARIS MONTPARNASSE"),
       aes(x=str_wrap(arrival_station,20),y= avg_delay_all_departing,color=month))+
       geom_jitter()+coord_flip()+ labs(color="Month")+
       transition_time(year)+ease_aes("linear")+
       geom_hline(yintercept = 0,color="red")+
       xlab("Arrival Station")+ylab("Average Delay All Departing")+
       ggtitle("Paris Montparnasse and its arrival Station" ,subtitle ="Year :{frame_time}")

animate(p,nframes=4,fps=1)
```

# Chosen City with Average Delay with All Arriving

```{r Paris Montparnasse avg delay all arriving}
p<-ggplot(subset(small_trains,departure_station=="PARIS MONTPARNASSE"),
       aes(x=str_wrap(arrival_station,20),y=avg_delay_all_arriving,color=month))+
       geom_jitter()+coord_flip()+ labs(color="Month")+
       transition_time(year)+ease_aes("linear")+
       geom_hline(yintercept = 0,color="red")+
       xlab("Arrival Station")+ylab("Average Delay All Arriving")+
       ggtitle("Paris Montparnasse and its arrival Station" ,subtitle ="Year :{frame_time}")

animate(p,nframes=4,fps=1)
```

# Chosen City with Number of Late Departures

```{r Paris Montparnasse num late at departure}
p<-ggplot(subset(small_trains,departure_station=="PARIS MONTPARNASSE"),
       aes(x=str_wrap(arrival_station,20),y=num_late_at_departure,
           color=month))+
       geom_jitter()+coord_flip()+ labs(color="Month")+
       transition_time(year)+ease_aes("linear")+
       xlab("Arrival Station")+ylab("Number of Lates at Departure")+
       ggtitle("Paris Montparnasse and its arrival Station" ,subtitle ="Year :{frame_time}")

animate(p,nframes=4,fps=1)
```

# Chosen City with Number of Late Arrivals

```{r Paris Montparnasse num arriving late}
p<-ggplot(subset(small_trains,departure_station=="PARIS MONTPARNASSE"),
       aes(x=str_wrap(arrival_station,20),y=num_arriving_late,
           color=month))+
       geom_jitter()+coord_flip()+ labs(color="Month")+
       transition_time(year)+ease_aes("linear")+
       xlab("Arrival Station")+ylab("Number of Lates at Arriving")+
       ggtitle("Paris Montparnasse and its arrival Station" ,subtitle ="Year :{frame_time}")

animate(p,nframes=4,fps=1)
```

# Departure Station with Average Journey Time and Total Number of Trips

```{r departure station with journey time avg and total num trips}
p<-ggplot(small_trains,aes(x=journey_time_avg,y=total_num_trips,color=month))+
      geom_point()+transition_states(departure_station)+labs(color="Month")+
      ggtitle("Average Journey Time and Total Number of Trips",
              subtitle="Departure Station : {closest_state}")+
      scale_y_continuous(breaks=seq(0,900,50),labels=seq(0,900,50))+
      scale_x_continuous(breaks=seq(0,500,50),labels=seq(0,500,50))+  
      xlab("Average Journey Time")+ylab("Total Number of Trips")+
      shadow_mark()

animate(p,nframes=59,fps=1)
```

# Departure Station with Average Delay All Departing and Number of Late at Departures

```{r departure station with Number of late and average Delay}
p<-ggplot(small_trains,aes(x=num_late_at_departure,y=avg_delay_all_departing,
                           color=month))+
      geom_point()+transition_states(departure_station)+labs(color="Month")+
      ggtitle("Average Delay at All Departing and Number of Lates at Departure",
              subtitle="Departure Station : {closest_state}")+
      geom_vline(xintercept = 0,color="red")+
      geom_hline(yintercept = 0,color="red")+
      scale_y_continuous(breaks=seq(-5,175,5),labels=seq(-5,175,5))+
      scale_x_continuous(breaks=seq(0,500,50),labels=seq(0,500,50))+  
      xlab("Number of Lates at Departure")+ylab("Average Delays at all Departing")+
      shadow_mark()

animate(p,nframes=59,fps=1)
```

# Departure Station with Average Delay All Arriving and Number of Arriving Late

```{r departure station with Number of Arriving late and average Delay arriving}
p<-ggplot(small_trains,aes(x=num_arriving_late,y=avg_delay_all_arriving,
                           color=month))+
      geom_point()+transition_states(departure_station)+labs(color="Month")+
      ggtitle("Average Delay at All Arriving and Number of Lates at Arriving",
              subtitle="Departure Station : {closest_state}")+
      geom_vline(xintercept = 0,color="red")+
      geom_hline(yintercept = 0,color="red")+
      scale_y_continuous(breaks=seq(-150,40,5),labels=seq(-150,40,5))+
      scale_x_continuous(breaks=seq(0,250,25),labels=seq(0,250,25))+  
      xlab("Number of Lates at Arriving")+ylab("Average Delays at all Arriving")+
      shadow_mark()

animate(p,nframes=59,fps=1)
```

# Delayed Cause and Delayed Number 

```{r Delayed No and Delayed cause}
small_trains %>%
    mutate(delay_cause = str_remove(delay_cause,"delay_cause_")) %>%
ggplot(.,aes(x=delay_cause,y=delayed_number))+
      xlab("Delay Cause")+ylab("Delayed Number")+
      ggtitle("Delayed Causes and Delayed Number as percentage")+
      geom_jitter()+coord_flip()
```

*THANK YOU*